### 2023-08-23 README.Rmd
update_230823_02

The goal of validateHOT is to validate the results of your validation
task (also known as holdout task). A validation task is essential to
make sure that your collected data of a MaxDiff, CBC, or ACBC are valid
and can also predict outside task that were not included in estimating
your utility scores. Although commercial studies often do not include a
validation/ holdout task (Yang et al., 2018), it is highly recommended to do so
(Orme, 2015; Rao, 2014). This validation/ holdout task does not only
help to check whether everything went right during data collection but
also to determine your final model. <code>validatHOT</code> provides
some of the relevant metrics to test the performance of your data in
predicting a holdout task. In [Sawtooth
Software's](https://sawtoothsoftware.com/) CBC a fixed validation/
holdout task is automatically implemented. If you conduct a MaxDiff or
ACBC study these have to be programmed by yourself.

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data,
and running your initial Hierarchical Bayes models, you can turn to
<code>validateHOT</code> and test how good your model predicts choices
in the validation/ holdout task. Herefore, you only have to insert your
raw utility scores as well as the actual choice of your validation/
holdout task. You can use the <code>merge()</code> provided by
<code>base</code> package (R Core Team, 2023). Afterward, you can read in your data
file and enjoy <code>validateHOT</code>.

@@

For all 3 key areas, the <code>createHOT</code> function is essential.
This function creates the total utilities for each alternative in the
validation/ holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. <code>createHOT</code> will do exactly this for you.

@@

<li><code>hitrate</code>: creates the Hit Rate (correctly predicted
choices) of your validation task. The output will contain the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage.
The number of correctly predicted participants' choices as well as the
percentage of how many choices were correctly predicted. If you specify
an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>kl</code>: Kullback-Leibler-Divergence which measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides both
divergence between predicted from observed and observed from predicted
due to the asymmetry of the Kullback-Leibler divergence. If you specify
an optional <code>group</code> argument the output is split by
groups. Currently, you can choose between \eqn{log{_2}} and \eqn{log} as logarithm base.</li>

<li><code>mae</code>: average absolute error, i.e., deviation between
predicted and stated choice share. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>medae</code>: since the averaged absolute error can be highly
influenced by the If you specify an optional <code>group</code> argument
the output is split by groups.</li>

<li><code>mhp</code>: averaged hit probability of participant's actual
choice in the validation/ holdout task. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>rmse</code>: provides the root-mean-squared error of deviation
between predicted and stated choice share. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

@@

@@

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the 5 provided functions, you
currently have to have a **none** option in your data. We currently
predict, e.g., whether a buy or no-buy was correctly predicted.
Information could be used for overestimating and underestimating,
respectively, of product purchases. In the following <code>TP</code>
stands for true positives, <code>FP</code> for false positives,
<code>TN</code> for true negatives, and <code>FN</code> for false
negatives.

@@

<li><code>accuracy</code>: defined as
$\frac{TP + TN}{TP + FP + TN + FN}$. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>f1</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>precision</code>: defined as $\frac{TP}{TP + FP}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>recall</code>: defined as $\frac{TP}{TP + FN}$. If you specify
an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>specificity</code>: defined as $\frac{TN}{TN + FP}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

@@

@@

<li><code>freqassort</code>: Inspired by the former <a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>freqassort</code>
will give you the averaged frequency, how many products the participants
will choose from your in the function determined potential assortment.
Again, you have to define a `none` alternative. `freqassort` uses the *threshold* approach,
meaning if the utility of one product is above the utility of `none`,
it is marked as potential purchase option. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

@@

<li><code>shareofpref</code>: provides you the aggregated share of
preference, including the lower and upper confidence interval, which is
calculated according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$. If
you specify an optional <code>group</code> argument the output is split
by groups and provided in a list element.</li>

@@

We are teaching a preference measurement seminar for students. Often
these students did not have experience with *R* before or only sparsely.
We teach them to also validate their results and wanted to give them an
easy way on how to do this in *R*. Of course, there are other great
packages which are faster in running (i.e., Hamner & Frasco, 2018), however,
these packages need some more data wrangling in order to use the
appropriate functions, which might be a burden or barrier for the one or
the other.

Moreover, as Yang et al. (2018) also report, often commercial studies do not use
any validation task. Some market research companies (as far as we have
realized) are not that familiar with *R* yet. Since these functions are
not always implemented in other softwares, this might be one reason why
they do not include one simply because they do not know how to use it
correctly. Having a package to evaluate the validation/ holdout task can
also be beneficial for those companies.

@@

> &#10071; <code>createHOT</code> is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by using the <code>names()</code> function or by using <code>which()</code> and <code>colnames()</code>, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of <code>id</code>, we could also use <code>which(colnames(CBC) == "ID")</code>.

```{r}
HOT <- createHOT(
  data = CBC,
  id = 1,
  None = 21,
  prod = 3,
  prod.levels = list(c(4, 9, 19), c(8, 12, 17), c(5, 10, 17)),
  coding = c(0, 0, 0),
  method = "CBC",
  choice = 22
)
```

Let us take a glimpse at the output, which shows the total raw utilities for each of the function included in the validation/ holdout task.

@@

In the next step, we would like to see how well our model (from which we took the raw utilities) predict the choices in the validation/ holdout task. First, we will test the <code>hitrate()</code> function. We specify the <code>data</code>, the column names of the alternatives (<code>opts</code>; remember we have three alternatives + the *no-buy* alternative), and finally the choice. We can see that the hitrate of our example is `r round(as.numeric(hitrate(data = HOT, opts = c(2:5), choice = choice)[3, 2]), digits  = 3)`.

```{r}
hitrate(data = HOT, opts = c(2:5), choice = choice)
```

@@

```{r}
mae(data = HOT, opts = c(2:5), choice = choice)
```
To cover also one example on how to use the metrics of the confusion matrix, we could test whether our model overestimates the purchase behavior (our model predicts a *buy* although participant opts for a *no* buy) or underestimates it (i.e., model predicts *no-buy* but participant opts for a *buy*). We will test the accuracy of the model by running the <code>accuracy()</code> function.

```{r}
accuracy(data = HOT, opts = c(2:5), choice = choice, none = None)
```
Finally, let us test, how many participants would at least buy one of the three products, assuming that this is one potential assortment we would like to offer to our consumers. We will use the <code>reach()</code> function and use the threshold approach. To specify the bundles we are offering we use the <code>opts</code> argument in our function.

```{r}
reach(data = HOT, opts = c(Option_1:Option_3), none = None)
```

@@

In a second example, we again use a *CBC*, however, this time we show how to use <code>validateHOT</code> if one of the variables are linear coded. All other examples are provided in the Vignette.

@@

Next, we create the validation/ holdout task to evaluate it in the next step. We use a validation/ holdout task with three alternatives plus the *no-buy* option, just as we did in the previous example. The only difference to the previous example is that we coded the third attribute (<code>Att3_Lin</code>) as linear.

@@

To interpolate the value, we have to provide <code>validateHOT</code> the <code>interpolate.levels</code>. These **need** to be the same as provided to, e.g., Sawtooth as levels. Moreover, it is important that the value that should be interpolated needs to lie within the lower and upper bound of <code>interpolate.levels</code>. In our case, we had 7 levels that range from 10 to 70. The value we want to interpolate for the second alternative is 40, which of course lies within 10 and 70.

Next, we define the column index of the linear coded variable (<code>lin.p</code>) and specify the <code>coding</code> we talked about above. Again, we are running a *CBC* specified by the <code>method</code> argument. This time, we would like to keep some of the variables to the data frame, which we specify in <code>varskeep</code>. We only keep one further variable, however, you can specify as many as you want. This could be relevant if you would like to display results per group. Finally, we just tell <code>validateHOT</code> the column index of the final choice (<code>choice</code>) and we are good to go.

```{r}
CBC <- createHOT(
  data = CBC_lin,
  id = 1,
  None = 15,
  prod = 3,
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)),
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)),
  lin.p = 14,
  coding = c(0, 0, 1),
  method = "CBC",
  varskeep = 17,
  choice = 16
)
```

@@

```{r}
hitrate(data = CBC, opts = c(Option_1:None), choice = choice, group = Group)
```

In this case, the Grouping variable is just an integer. However, the output is different if it is a factor or if it is labelled data. To proof this we just quickly change <code>group</code> in a factor by using the <code>factor()</code> function provided by R Core Team (2023).

```{r}
CBC$Group <- base::factor(CBC$Group,
  levels = c(1:3),
  labels = paste0("Group_", c(1:3))
)
```

@@

```{r}
mhp(data = CBC, opts = c(Option_1:None), choice = choice, group = Group)
```

@@

>>>>

The goal of validateHOT is to validate the results of your validation
task (also known as holdout task). A validation task is essential to
make sure that your collected data of a MaxDiff, CBC, or ACBC are valid
and can also predict outside tasks that were not included in estimating
your utility scores. Although commercial studies often do not include a
validation/holdout task (Yang et al., 2018), it is highly recommended to do so
(Orme, 2015; Rao, 2014). This validation/holdout task does not only
help to check whether everything went right during data collection but
also to determine your final model. <code>validatHOT</code> provides
some of the relevant metrics to test the performance of your data in
predicting a holdout task. In [Sawtooth
Software's](https://sawtoothsoftware.com/) CBC a fixed validation/
holdout task is automatically implemented. If you conduct a MaxDiff or
ACBC study, you have to program these on your own.

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data,
and running your initial Hierarchical Bayes models, you can turn to
<code>validateHOT</code> and test how good your model predicts choices
in the validation/holdout task. Herefore, you only have to read in your
raw utility scores as well as the actual choice of your validation/
holdout task. You can use the <code>merge()</code> provided by the
<code>base</code> package (R Core Team, 2023). Afterward, you can read in your data
file and enjoy <code>validateHOT</code>. We provide a tutorial a short tutorial in the vignette.

@@

For all 3 key areas, the <code>createHOT</code> function is essential.
This function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. <code>createHOT</code> will do exactly this for you.

@@

<li><code>hitrate</code>: creates the Hit Rate (correctly predicted
choices) of your validation task. The output will contain 
the percentage of how many choices were correctly predicted, 
the chance level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the number of observations. 
</li>

<li><code>kl</code>: Kullback-Leibler-Divergence measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides both
divergence between predicted from observed and observed from predicted
due to the asymmetry of the Kullback-Leibler divergence. If you specify
an optional <code>group</code> argument the output is split by
groups. Currently, you can choose between \eqn{log} and \eqn{log{_2}} as logarithm base.
The default is set to \eqn{log}.</li>

<li><code>mae</code>: mean absolute error, i.e., deviation between
predicted and stated choice share. </li>

<li><code>medae</code>: median absolute error, which is less affected
by outliers compared to the mean absolute error.</li>

<li><code>mhp</code>: averaged hit probability of participant's actual
choice in the validation/holdout task.</li>

<li><code>rmse</code>: provides the root-mean-square error of deviation
between predicted and stated choice share.</li>

@@

All functions can be extended with the `group` argument to get output by group.

@@

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the 5 provided functions, you
currently have to include a **none** alternative in your validation/holdout task. We currently
predict, e.g., whether a buy or no-buy was correctly predicted.
Information could be used for overestimating and underestimating,
respectively, of product purchases. In the following <code>TP</code>
stands for true positives, <code>FP</code> for false positives,
<code>TN</code> for true negatives, and <code>FN</code> for false
negatives (Burger, 2018). 
To translate this to the `validateHOT` logic, imagine you have a validation/holdout task
with 5 alternatives plus the alternative not to buy any of those chosen. `validateHOT` now measures
whether or not a buy (participant opts for one of the 5 alternatives) or a no-buy (participant opts
for the no-buy alternative), respectively, is correctly predicted.

@@

<li><code>accuracy</code>: calculates the number of correctly predicted choices (buy or no-buy);
$\frac{TP + TN}{TP + TN + FP + FN}$ (Burger, 2018). 
</li>

<li><code>f1</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$.</li>

<li><code>precision</code>: defined as $\frac{TP}{TP + FP}$ (Burger, 2018).</li>

<li><code>recall</code>: defined as $\frac{TP}{TP + FN}$ (Burger, 2018).</li>

<li><code>specificity</code>: defined as $\frac{TN}{TN + FP}$ (Burger, 2018).</li>

@@

Again, all functions can be extended with the `group` argument to get output by group.

@@

<li><code>freqassort</code>: Inspired by the former <a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>freqassort</code>
will give you the averaged frequency, how many products the participants
will choose from your in the function determined potential assortment.
Again, you have to define a `none` alternative, because `freqassort` uses the *threshold* approach,
meaning if the utility of one product is above the utility of `none`,
it is marked as potential purchase option.</li>

@@

<li><code>shareofpref</code>: provides you the aggregated share of
preference, including the standard error, as well as the lower and upper confidence interval, which is
calculated according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$ (Orme, 2020, p. 94).</li>

@@

We are teaching a preference measurement seminar for students. Often
these students did not have any prior experience (or only sparsely) with *R*.
One of the chapters in this class is about model validating the results obtained and we teach this,
of course, in *R*.
Of course, there are other great packages which are faster in running (i.e., `Metrics` by Hamner & Frasco, 2018), 
however, these packages need some more data wrangling in order to use the
appropriate functions, which might be a burden or barrier for one or
the other.

Moreover, as Yang et al. (2018) report, commercial studies often do not use
any validation task. Again, the missing experience in *R* could be one explanation. 
Since these functions are not always implemented in other softwares, this might be one reason why
they do not include one simply because they do not know how to use it
correctly. Having a package to evaluate the validation/holdout task can
also be beneficial from this perspective.

@@

> &#10071; <code>createHOT</code> is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by using the <code>names()</code> function or by using <code>which()</code> and <code>colnames()</code>, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of <code>ID</code>, we could also use <code>which(colnames(CBC) == "ID")</code>.

```{r}
HOT <- createHOT(
  data = CBC, # data frame
  id = 1, # column index of the id
  None = 21, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 19), c(8, 12, 17), c(5, 10, 17)), # column index of the attribute levels for each alternative
  coding = c(0, 0, 0), # how the attributes were coded
  method = "CBC", # method
  choice = 22 # column index of choice
)
```

Let us take a glimpse at the output, which shows the total raw utilities for each of the alternatives 
included in the validation/holdout task.

@@

In the next step, we would like to see how well our model (from which we took the raw utilities) predict 
the actual choices in the validation/holdout task. First, we will test the <code>hitrate()</code> function. 
We specify the <code>data</code>, the column names of the alternatives (<code>opts</code>; remember 
we have three alternatives + the *no-buy* alternative), and finally the actual choice.

```{r}
hitrate(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```

@@

```{r}
mae(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```
To cover also one example on how to use the metrics of the confusion matrix, we could test whether our model overestimates the purchase behavior (our model predicts a *buy* although participant opts for a *no* buy) or underestimates it (i.e., model predicts *no-buy* but participant opts for a *buy*). We will test the accuracy of the model by running the <code>accuracy()</code> function.

```{r}
accuracy(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  none = None # column name of none alternative
)
```
Finally, let us test, how many participants would at least buy one of the three products, assuming that this is one potential assortment we would like to offer to our consumers. We will use the <code>reach()</code> function and use the threshold approach. To specify the bundles we are offering we use the <code>opts</code> argument in our function.

```{r}
reach(
  data = HOT, # data frame
  opts = c(Option_1:Option_3), # products that should be considered
  none = None # column name of none alternative
)
```

@@

In a second example, we again use a *CBC*, however, this time we show how to use <code>validateHOT</code> if one of the variables are linear coded. All other examples are provided in the accompanied vignette.

@@

Next, we create the validation/holdout task to evaluate it in the next step. We use a validation/holdout task with three alternatives plus the *no-buy* alternative, just as we did in the previous example. The only difference to the previous example is that for our model estimation, the third attribute (<code>Att3_Lin</code>) was coded as linear.

@@

To interpolate the value, we have to provide <code>validateHOT</code> the <code>interpolate.levels</code>. These **need** to be the same as provided to Sawtooth as levels. Moreover, it is important that the value that should be interpolated needs to lie within the lower and upper bound of <code>interpolate.levels</code>. In our case, we had 7 levels that range from 10 to 70.

Next, we define the column index of the linear coded variable (<code>lin.p</code>) and specify the <code>coding</code> we talked about above. Again, we are running a *CBC* specified by the <code>method</code> argument. This time, we would like to keep some of the variables in the data frame, which we specify in <code>varskeep</code>. We only keep one further variable, however, you can specify as many as you want. This could be relevant if you would like to display results per group. Finally, we just tell <code>validateHOT</code> the column index of the final choice (<code>choice</code>) and we are all set.

```{r}
CBC <- createHOT(
  data = CBC_lin, # data frame
  id = 1, # column index of the id
  None = 15, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)), # column index of the attribute levels for each alternative (for linear coded we specify value to be interpolated)
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)), # actual values for the levels that should be interpolated
  lin.p = 14, # column index of the linear coded variable
  coding = c(0, 0, 1), # coding of the 3 attributes; 0 = part-worth, 1 = linear, 2 = piecewise
  method = "CBC", # method
  varskeep = 17, # column index of variables that should be kept in the dataframe
  choice = 16 # column index of choice
)
```

@@

```{r}
hitrate(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

In this case, the group assignment is stored in form of integers. However, the output is the same if it is a factor or if it is labelled data. To proof this we just quickly change <code>group</code> into a factor by using the <code>factor()</code> function provided by R Core Team (2023).

```{r}
CBC$Group <- base::factor(
  CBC$Group,
  levels = c(1:3),
  labels = paste0("Group_", c(1:3))
)
```

@@

```{r}
mhp(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

@@

Orme, B. K. (2020). <em>Getting Started with Conjoint Analysis: Strategies for Product Design and Pricing Research</em>. 4th edition. Manhattan Beach, CA: Research Publishers LLC.
